### 2013-09-26 docs/proposal/doc.tex
Small revisions to final sections. This version was sent for feedback.

% Needed packages
\documentclass[a4paper, 10pt, english, twocolumn]{article}
\usepackage[english]{babel}
\usepackage[cm]{fullpage}
\usepackage{cite}
\usepackage{anysize}
\usepackage[compact]{titlesec}
\usepackage{graphicx}
\usepackage{stfloats}
\usepackage{listings}
\usepackage{hyperref}

\usepackage{amssymb,amsmath}
\usepackage{algorithmicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}

% for coloring individual cells in a table
\usepackage[table]{xcolor}
%\usepackage{pgfgantt}

\newcommand{\keywords}[1]{\par\noindent 
{\bf Keywords\/}. #1}

% Margins & Headers
\marginsize{2.5cm}{2.5cm}{2cm}{1.5cm}
\columnsep 0.4in
\footskip 0.4in 
\usepackage{changepage}

% E-mail formatting
\usepackage{color,hyperref}
    \catcode`\_=11\relax
    \newcommand\email[1]{\_email #1\q_nil}
    \def\_email#1@#2\q_nil{
      \href{mailto:#1@#2}{{\emailfont #1\emailampersat #2}}
    }
    \newcommand\emailfont{\sffamily}
    \newcommand\emailampersat{{\color{red}\small@}}
    \catcode`\_=8\relax

% List modifications
\newenvironment{packed_item}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\newenvironment{packed_enum}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

% ### Mathematics ###
\newcommand{\bpm}{\begin{pmatrix}}
\newcommand{\epm}{\end{pmatrix}}

\newcommand{\bbm}{\begin{bmatrix}}
\newcommand{\ebm}{\end{bmatrix}}

\newcommand{\bsm}{\bigl( \begin{smallmatrix}}
\newcommand{\esm}{ \end{smallmatrix} \bigl)}

\newcommand{\mbf}{\mathbf}

% ### Matrices and Vectors ###
\newcommand{\mtx}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand*\Let[2]{\State #1 $\gets$ #2}

% ### Sets ###
\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}

% ### Other ###

\newcommand{\transpose}{^{T}}
\newcommand{\inv}{^{-1}}
\newcommand{\pseudoinv}{^{+}}
% ############## End Macros ##############

% Title
\title{\fontfamily{phv}\selectfont{Causal Discovery methods for Effective Connectivity}}
\author{
  \textbf{Ramon Janssen} - \href{mailto:ramon.janssen@student.ru.nl}{ramon.janssen@student.ru.nl} \\
  \textbf{Tom de Ruijter} - \href{mailto:t.deruijter@student.ru.nl}{t.deruijter@student.ru.nl}
}

\date{\fontfamily{ptm}\selectfont{\small{\bfseries{\today - Radboud
Universiteit Nijmegen}}}\\[0.5cm]\rule{\linewidth}{0.3mm}}

\begin{document}

\maketitle

\begin{abstract}
Lorem Ipsum Dolor Sit Amet.
\keywords{Keywords}
\end{abstract}

\setlength{\parindent}{0.0cm}
\setlength{\parskip}{0.25cm}

\section{Introduction}

% vergelijking met granger causality, en andere methodes?
% granger (Alard Roebroeck, Elia Formisano, Rainer Goebel, 2005) -> geen structural connectivity nodig (schaalt niet goed)
% covariance structural equation modeling -> structural connectivity nodig (Roebroeck, Formisano, Goebel, 2005)
% Multivariate autoregressive modeling (Harrison, Penny, Friston) -> Set of interaccting regions chosen beforehand (Roebroeck, Formisano, Goebel, 2005)
% Dynamic causal modelling (Friston, Harrison, Penny, 2003) -> Set of interaccting regions chosen beforehand (Roebroeck, Formisano, Goebel, 2005)

\paragraph{Motivation}
The brain and particularly the human brain have been studied for hundreds of years.
Today, the secrets of our brains still are one of the most sought-after.
The techniques however have changed.
Today, advanced measurement methods exist that allow scientists to peek inside brains that give a coarse but broad overview without opening a single skull.

Of particular interest are causal relations between brain regions and cognitive and bodily functions.
We have come a long way since the time of Phrenology, the pseudoscience based attempting to derive cognitive ability and personality from measurements of the skull or, post-mortem, the brain.
The question whether brain region X is somehow connected to brain region Y is still highly relevant; it is the main question of the present-day scientific field of brain connectivity.
Understanding brain structure implies understanding more about the brain as a whole.

Not surprisingly, brain connectivity also finds applications in medicine.
Neuro-degenerative diseases such as Alzheimer's disease, Parkinson's disease, dementia, Amyotrophic lateral sclerosis (ALS) have all been shown to severely alter brain connectivity. %TODO: Citation needed.
Better methods for analysing connectivity could lead to more insight in these diseases.

The field of brain connectivity finds its roots in the early 1990s \cite{friston1993functional, friston1994}, though because of more recent developments in the field of Artificial Intelligence and Machine Learning it is possible to do more thorough analysis \cite{vandenheuvel2010}.
One type of connectivity analysis involves finding causal relations between brain areas, specifically \emph{effective connectivity}.
However, causal analysis is also relatively new and its application in brain connectivity analysis still faces many challenges that require further research \cite{ramsey2010}.

\paragraph{Problem statement}
This project seeks to combine the strengths of current state-of-the-art causal discovery methods and apply them to resting state brain fMRI data in order to find new applicable methods for determining causal patterns in the human brain.

The work by \cite{ramsey2010} addresses several problems causal analysis in fMRI analysis suffers from.
Methods in causal discovery have high computational bounds, making it nearly impossible to calculate networks with more than hundreds of nodes.
Another import fact is fMRI Analysis indirectly measures brain activity.
Brain models that do not account for possible latent - indirect - sources within the brain or the shortcomings of fMRI may suffer from noise or fail to capture underlying patterns.
Another problem is the strong diversity between measurement subjects.
Every brain is inherently different, making it non-trivial to combine subject data or even draw conclusions across subjects.
Even within a single brain, how different areas react over time changes.

Several methods for finding effective connectivity already exist \cite{mclntosh1994, harrison2003, friston2003, roebroeck2005}.
However using more generally applicable causal discovery methods has not been elaborated on much.
Also, none of these provide a measure of uncertainty.
It is a fact that there are errors in any graph produced.
This is inherent to the methods and the coarseness of the measurements.
As there is no standard baseline to compare results with, a measure of uncertainty would be highly preferable.
A framework for such a probabilistic approach is introduced in \cite{claassen2012}.

We would like to know whether applying such methods, such as the approach introduced in \cite{claassen2012}, or a standard approach such as PC-algorithm \cite{spirtes2000} can solve some of the difficulties effective connectivity suffers from.

% Structure of the rest of proposal
In the remainder of this proposal we briefly discuss some necessary background knowledge on connectivity and methods, we then discuss our strategy and finally we present a time-plan overview.

\section{Theory}

\subsection*{Brain connectivity}

%Question: should this be included?
\subsubsection*{Functional connectivity}
Functional connectivity describes the statistical dependance of neuronal activity between different brain regions \cite{friston1993functional}.
It gives an insight in the organisation of the brain, even though functioning is not immediately visible.
Functional connectivity is strongly time-dependant, as activity changes rapidly providing only a small time window to measure.
This means measured dependencies can be the result of statistical noise.
Functional connectivity can be deduced by fMRI measurement of brains in resting-state \cite{Lowe2000, doria2010, Bullmore2009}.
Resting state indicates that subjects are instructed to relax without thinking of anything in particular to stimulate spontaneous brain activity.

\subsubsection*{Structural connectivity}
Structural connectivity is defined as the mapping of anatomical - neural - paths in the brain between different brain regions \cite{friston1994}.
This is strongly related with functional connectivity: regions can only be structurally connected if there is a functional relation between them.
This is intuitive relation can be demonstrated empirically \cite{vandenheuvel2009}.
The other way around it can be demonstrated that functional connectivity is constrained by structural connectivity \cite{cabral2012}.
Structural connectivity is less time dependent as it involves mappings of anatomical connections opposing temporal activity patterns.

\subsubsection*{Effective connectivity}
In contrast to functional and structural connectivity, effective connectivity takes into account the cause and effect of relations.
It has been described as ``the influence one neural system exerts over another'' \cite{friston1994}.
Effective connectivity indicates which brain regions stimulate other regions.
Time series need to be analysed to infer effective connectivity as cause and effect can be deduced from which event precedes which. 
It is possible to infer effective connectivity from structural and functional connectivity \cite{mclntosh1994, harrison2003, friston2003, roebroeck2005}.

Research has been done on inferring effective connectivity directly with several methods.
The most renowned methods are Granger causality \cite{roebroeck2005}, Structural Equation Modelling \cite{mclntosh1994}, Multivariate Autoregressive Modeling \cite{harrison2003} and Dynamic Causal Modelling \cite{friston2003}.
These are all methods dedicated to this problem and no attempt has been done from the viewpoint of causal discovery in graphs.
These methods also suffer from similar problems of which perhaps the most important is scaling.
Most of these methods can only handle tens of nodes in reasonable time.

\subsection*{Causal discovery methods}
The first causal method we will apply on brain structure data will be the PC-algorithm \cite{spirtes2000}.
It is used to find structure, as well as causal relations between vertices in directed graphs. 
If we already provide a graph skeleton, PC-algorithm can be used to determine directions.
An adaptation of PC that is better adapted to finding latent variables is Fast Causal Inference \cite{spirtes2000}, though this method does not scale well to tens of variables.

Another causal discovery method is the BCCD-algorithm \cite{claassen2012}, which gives similar results as PC.
This algorithm has some useful properties; it is a Bayesian approach and as such, it is more robust against noise.
The authors have shown the method to perform as good as or better than other causal discovery algorithms such as PC on representative datasets.
As is inherent to Bayesian methods, or approximations of Bayesian methods, the resulting network graph is not a  single structure but rather a distribution over possible structures.
Concretely, this provides an uncertainty measure over edges in the network, giving additional information compared to PC.

\begin{figure*}[bp]
\begin{tabular}{c || c | c | c | c | c | c | c | c | c | c |}
\setlength{\tabcolsep}{12pt}
\renewcommand{\arraystretch}{1.3}
Week          & 41 & 42 &  43 &  44 &  45 &  46 &  47 &  48 &  49 &  50 \tabularnewline \hline
PC            & \multicolumn{2}{ |c| }{\cellcolor[gray]{0.9} Find implementation} & \multicolumn{2}{ |c| }{\cellcolor[gray]{0.9} Structural}  & \cellcolor[gray]{0.9} Functional & \multicolumn{4}{ |c| }{} & \tabularnewline \hline
BCCD          & \multicolumn{4}{ |c| }{} & \multicolumn{2}{ |c| }{\cellcolor[gray]{0.85} Structural} & \cellcolor[gray]{0.85} Functional & \multicolumn{3}{ |c| }{} \tabularnewline \hline
Analysis      & \multicolumn{3}{ |c| }{} & \multicolumn{2}{ |c| }{\cellcolor[gray]{0.9} PC} & & \multicolumn{2}{ |c| }{\cellcolor[gray]{0.9} BCCD} & \multicolumn{2}{ |c| }{} \tabularnewline \hline
Write article & \multicolumn{3}{ |c| }{} & \multicolumn{7}{ |c| }{\cellcolor[gray]{0.85}}  \tabularnewline \hline
\end{tabular}
\caption{Project time table. `Structural' implies applying the given method on structural data only. `Functional' implies inferring structure with the given method first through functional data.}
\end{figure*}

\subsection*{Data acquisition \& analysis}
We will apply the mentioned methods on structural and functional data gathered from human subjects.
We use the same functional and structural data as in Hinne et al. \cite{hinne2013}, of which we have selected the best 6 subjects based on MAP results.
To find effective connectivity, both are needed.
Structural connectivity of these subjects follows directly from DTI-measurements.
The functional connectivity can be extracted from these fMRI-measurement, possibly in combination with the structural connectivity.
This data is already available.

\section{Methods \& Strategy}
%Short general introduction to both methods
We use existing implementations of PC and BCCD to infer effective connectivity.
An implementation of BCCD suitable for this project is made available by the department of Intelligent Systems at the Radboud University Nijmegen.

The first step of this step will be to get the implementations for the algorithms working on the datasets with functional and structural data.

Firstly we will use the PC-algorithm on given structural networks to infer causal relations.
Later we also infer network structure with the PC-algorithm to compare with the previous.
A suitable implementation for PC-algorithm has to be found and selected.

The second step is to apply the BCCD-algorithm, also to be used with and without provided network structure.
Both methods provide a connectivity matrix representing the effective connectivity.

Results will then be evaluated and compared over all six subjects to see whether gained data is consistent.
If so, we will analyse the properties and difficulties presented by Ramsey et al. \cite{ramsey2010}.
We will conclude whether standard causal discovery methods provide a good direction of research for deriving and evaluating effective connectivity.
In any case, there is enough data available to achieve some useful results to draw conclusions.

% Licht hier toe met welke methoden we gaan vergelijken en wat we nog meer besproken hebben in de outline.
% - Difficulties in causal discovery
% - Comparison methods
% - Comparison procedure
% - Properties

% week 50 is deadline voor draft, 51 is presentaties, 2 inleveren van final
% gantt gaf wat problemen met packages die nodig zijn om te compileren
% \section{Time schedule}
% \begin{ganttchart}{1}{10}
% \gantttitle{2011}{10} \\
% \gantttitlelist{40,...,52,1,2}{1} \\
% \ganttgroup{Group 1}{1}{7} \\
% \ganttbar{Task 1}{1}{2} \\
% \ganttlinkedbar{Task 2}{3}{7} \ganttnewline
% \ganttmilestone{Milestone}{7} \ganttnewline
% \ganttbar{Final Task}{8}{12}
% \ganttlink{elem2}{elem3}
% \ganttlink{elem3}{elem4}
% \end{ganttchart}

\section{time schedule}
This project is divided in four main parts: applying the PC-algorithm, applying the BCCD-algorithm, analysing the data (including cross-subject comparison) and writing the article. The schedule for applying the algorithm has been subdivided into applying it with functional connectivity data only, and applying it to both functional and structural connectivity data.

\bibliography{references}{}
\addcontentsline{toc}{section}{References}
\bibliographystyle{apalike}

\end{document}